Colonies - Compute Continuums across Platforms

### 2023-04-11 paper_2023/paper.tex
first draft

Today, utilization of serverless computing is experiencing a significant growth \cite{cognit}, primarily due to its potential to liberate developers from the burden of managing underlying cloud infrastructures. Attempts are currently being made to implement serverless workflow management systems. For example, OpenWolf \cite{openwolf} is a serverless workflow engine designed to utilize the Function-as-a-Service (FaaS) paradigm for composing complex scientific workflows. It is based on OpenFaaS \cite{openfaas}, which allows functions to run on Kubernetes clusters. 
The serverless workflow project \cite{serverlessworkflows} aims at providing a vendor-neutral workflow DSL. Synapse \cite{synapse} is a workflow management system similar to Argo, but that implements the serverless workflow specification. A similar engine is proposed in \cite{scafe}. 
Colonies can also be used to implement serverless workflow management systems. This will be further discussed in Section \ref{faas}. However, in contrast to previous work, Colonies is platform independent and does not require Kubernetes. By using a zero-trust security protocol, functions can be securely executed by distributed executors deployed anywhere on the Internet. It is important to point out that Colonies does not provide an infrastructure for function execution. Instead, Colonies primary role is to serve as a platform for coordinating function executions which are carried out by distributed executors.

@@

Although microservices have attractive properties and simplify both development and deployment of executors, they also introduce complexity to the Colonies servers. To implement a workflow management framework supporting distributed microservices, the following challenges must be addressed:

@@

\item \textbf{Service discovery:} The Colonies framework must enable executors to dynamically register and deregister to Colonies servers. It must be possible to deploy executors anywhere on the Internet, even behind firewalls.

@@

The Colonies framework uses a combination of technologies to address the aforementioned challenges. A fundamental design principle is statelessness, which makes Colonies simpler, more reliable and scalable, and easier to implement. To ensure reliability and data consistency, distributed consensus algorithms are needed to offer high-availability and handle server crashes. The remainder of this section describes how the Colonies framework is implemented. The next discusses the role of queues to support batch jobs and realtime processing, but also how queues can be used to enable loose coupling of system components.

@@

Table \ref{proctable} shows an example of a process table. The process table also contains a reference to a function specification depicted in Table \ref{functable}. To assign a process to an executor, the Colonies server searches in the process table to find a process matching a waiting executor. By using the \emph{priority time} column to sort the processes, the process table serves as a queue. This can be done in SQL by utilizing the \emph{order by} to sort processes according to their submission time. To make it possible to handle priorities, the submission time is adjusted so that higher priority processes are processed before lower priority processes. When a process is submitted, a \emph{priority time} value is calculated and stored in the process table. Equation \ref{eq:pt} shows how the priority time is calculated for a nanosecond timestamp.

@@

To minimize downtime, Colonies supports high-availability deployments. If one Coloniser server crashes, an executor simply need to resend the failed request, which will then be served by another Colonies server replica. However, by introducing multiple Colonies servers, there is again a risk of race conditions when assigning processes to executors. This means that the Colonies server replicas must coordinate which replica server incoming assign requests so that precisely one executor is assigned to a process.

Raft \cite{raft} is a consensus algorithm specifically designed to manage a replicated log within a distributed system. It functions within a cluster of servers, where a single server takes on the role of leader while the remaining servers act as followers. The leader is responsible for managing the replicated log, processing client requests, and replicating entries to the followers. Followers passively replicate the leader's log and participate in leader elections. The leadership can change over time due to elections triggered by timeouts or other factors.

Adding Raft with the Colonies framework allows incoming assign requests to be directed towards the leading Colonies server, thereby ensuring that only one Colonies server replica handles such requests. Figure \ref{fig:ha_deployment} shows an overview of a high-availability Colonies deployment. A new leader is elected in the event that a Colonies server replica fails. The Raft protocol also enables seamless updates to the Colonies server software by making it possible to upgrade each replica individually. Consequently, Colonies is well-suited for Kubernetes deployments, ensuring high levels of availability and fault tolerance.

@@

\subsubsection{Workflows}
A workflow is a series of processes that need to be completed in a specific order. In Colonies, workflows are represented as directed acyclic graphs (DAGs) where nodes represent processes and edges represent dependencies and data flow between processes. Similar to processes, workflows are managed completely stateless. When a workflow is submitted, all processes part of the workflow are submitted and added to the process table similar to how ordinary processes are handled. To control the order processes can be assigned, Colonies sets a flag, \emph{wait for parents} to prevent processes to be assigned to an executor before their dependencies have been satisfied. Note that processes may run in parallel if a sufficient number of executors are available.

@@

When a process is closed, the Colonies server checks its child processes to see if all their parent processes have also completed. If all parents of a child process have finished, the server updates the process table in the database by setting the flag \emph{wait for parents} to false, which allows child processes to be assigned to executors. This operation is also done statelessly when processing the close request.

@@

To be able to generate a DAG, an additional database table is required. Table \ref{deptable} shows an example of a dependency table storing relationships between processes. The table is updated when a workflow is submitted to a Colonies server. As mentioned before, when an executor is assigned a process it obtains exclusive rights to that process within the process table. This exclusivity also means that it is possible for an exector to dynamically add new children processesto  the workflow DAG on the fly. As a result, it becomes possible to implement MapReduce \cite{mapreduce} data processing patterns similar to Hadoop \cite{hadoop}.

@@

Cron is a time-based job scheduler designed to trigger workflows at predefined intervals. It can for example be used to automate tasks such as routinely fetching data from external systems. To make Colonies robust and easy to implement it is essential to make Cron workflows stateless. This can be achieved by introducing an additional database table to store information about Cron workflow, and let the elected Colonies server leader assume the responsibility of managing Cron workflows through a two-step process.

@@

\subsubsection{Generators}
Another feature of the Colonies framework are so-called generators. Generators are workflow templates that are automatically triggered when specific conditions are met. Generators can be used to accumulate input data into bulks to improve performance or parallelism. For example, an external service downloads satellite images and can use a generator to automatically triggered a workflow when 10 images are collected. This is achieved by sending a special \emph{pack request} containing input data to a Colonies server, targeting a certain generator.

@@

An advantage of using a stateless architecture for workflow management as proposed in the paper is the ability to monitor executor performance to detect any unusual or suspicious behaviors. Because all transactions are recorded in a database, it is possible to track the activity of individual executors and identify any unusual patterns or behaviors. For example, if an executor suddenly starts to consume a lot of resources, it could indicate a performance issue, or it could be a sign of malicious activity, such as a denial-of-service attack. Automatic monitoring can for example be implemented using anomaly detection algorithms and machine learning to grade the performance of each executor. Administrators can then be automatically notified if the grade of an executor exceeds a certain threshold, enabling them to take appropriate action.

@@

\subsection{Python SDK}
A colonies application typically consists of a network of executors written in different programming languages and deployed across multiple platforms. The executors are small microservices that interact with one another by submitting function specifications or workflows, which are then executed by other executors that are part of the same colony.

The first step in implementing an executor in Python is to create a private key and identity, followed by registering the executor with a colony. Registering an executor can only be carried out by the colony owner with access to the colony private key. Code in listening \ref{code:regexecutor} illustrates how to register an executor and which functions a specific executor is capable of running. In the example, an executor of type \emph{helloworld\_executor} is registered to run a function named \emph{helloworld}.

@@

\begin{figure}
     \centering
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[scale=0.17]{dashboard1.png}
         \caption{Screenshot of process information.}
     	 \label{fig:dashboard1}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[scale=0.17]{dashboard2.png}
      	 \caption{Screenshot of a workflow.}
     \end{subfigure}
     \caption{Screenshot of the Colonies Dashboard tool.}
\end{figure}

@@

Figure \ref{fig:dashboard1} depicts a screenshot of the Colonies dashboard tool, which enables users to monitor the status of processes and executors. The timeline depicted in the figure allows for real-time monitoring of the execution of processes, similar to tracking a physical package sent through a postal service. As previously discussed, it's also possible to subscribe to a process and receive information about its progress, which can be used as another point of integration. For example, sending a notification on Slack or by email if a process fails.

\subsection{Workflows}
A workflow is just a set of processes with dependencies represented as a DAG internally in Colonies. Workflows can either be submitted using SDKs or the Colonies CLI, which supports submission of JSON files. Listening \ref{code:workflow_spec} shows an example of a workflow consisting of four processes, each requiring different types of executors. The management of workflows is entirely handled by Colonies, and each executor executes only processes that are part of the graph for which it is responsible for. This approach significantly reduces the complexity of designing, implementing and operating large-scale AI applications and data pipelines.

@@

In addition to monitoring processes and executors, the Colonies dashboard can be used to visualize workflows. Figure \ref{fig:dashboard2} illustrates the DAG created when submitting the workflow shown in Listing \ref{code:workflow_spec}. This functionality makes it possible to follow workflow execution progress.

@@

\section{Use cases}
Colonies is currently used commerically by a startup company called RockSigma AB\cite{https://www.rocksigma.com} to implement a compute engine for analysing seismic activities in underground mines. This section will not dig into the details of this specific implementation; instead it will provide an outline of potential future use cases in which Colonies could make a considerable impact.

\subsection{Serverless computing}
\label{faas}
Serverless computing is a cloud-based architecture that allows developers to build and deploy applications without the need to manage the underlying infrastructure. In this model, the cloud service provider dynamically allocates and manages resources such as adding computing power or storage required by an application. Function-as-a-Service (FaaS) is a serverless computing model where developers provide a code snippet that should be executed, and the cloud provider takes care of scaling, configuration and deployment. FaaS is often used in event-driven architectures where functions are triggered by specific events, thus eliminating the need to keep servers continuously running. The Cognit project \cite{cognit} aims to extend the FaaS concept by supporting geographically distributed edge servers in addition to cloud resources. The project is also going to develop an AI-enabled orchestrator that can decide where to deploy FaaS code.

Colonies offer functionality to call functions in remote executors, a fundamental requirement to implement a FaaS platform. Colonies could also provide service discovery and provide security isolation to prevent non-authorized function calls. In some cases, it could make sense to manually manage a colony by an administrator, and manually registering and deploying executors. Alternatively, a software solution could be developed to automate the administration and automatically deploy or undeploy executors. Such an approach would be very similar to a Function-as-a-Service (FaaS) platform.

Figure \ref{fig:faas} depicts a Colonies-based FaaS platform. Similar to AWS Lambda or Google Cloud Functions, a developer would need to upload a code snippet, which would then be packaged and published as a container. To invoke a function, a function specification needs to be submitted  to the Colonies server indicating which function to call.

The advantage of Colonies is that it introduces an extra level of separation by offering a queuing system. If an executor is unavailable, the process is simply enqueued in the process table, enabling development of a loosely coupled system. A software agent could then monitor pending processes on the Colonies server, predict the system load, and automatically deploy executors to meet the load, resulting in a highly efficient and scalable system. In the future, the Cognit AI orchestrator could also be used together with Colonies to reduce energy consumption or minimize network latency by efficiently utilizing edge servers.

\subsection{HPC integration}
\label{hpc_use_case}
Slurm \cite{slurm} is a workload manager and job scheduler commonly used in HPC environments. In constrast to Colonies, it provides functionality to reserve resources. This is particularly useful when running complex simulations that requires a large number of of CPUs or a large amount of memory.

@@

Colonies could be used to fill the gap between cloud and HPC and create workflows that can automate processing of data in the cloud while also taking advantage of the vaste compuational capacity offered by HPC. For example, a company could develop a workflow that automatically copy training data to a HPC center, train a neural network, move the result back to the cloud, and finally re-deploy a Kubernetes service to use the new model.

Figure \ref{hpc_workflow} shows an example how Colonies could be used in a HPC environment. Several executors runs at the HPC system, each with their own responsibilities. The Container Importer Executor is responsible for converting container images to a format that is supported by Slurm, e.g. Singularity containers\footnote{https://docs.sylabs.io}. The Data Importer Executors downloads data to the HPC file system. The Slurm Executor is responsible for submitting and overseeing Slurm jobs, while the Data Exporter executor is responsible for uploading result data to the cloud.

A challenge developing cross-platform workflows is the integration with different autentication services. For example, the Data Importer may need to access an S3 services to fetch data. The Container Importer must authenticate with a Docker registry to be able to pull a Docker container to the HPC system. Generally, it is not straigtforward connect to the HPC system. In many cases, users have to manually set up SSH tunnels.

When using Colonies, users would not direclty connect to the HPC system, but indirectly control it by submitting Colonies processes that are subsequently picked up by HPC executors.

picked up the HPC executors. Since the security model offered by Colonies allows for different executors to trust each other, Colonies also provide a way of sharing secrets, e.g. S3 secrets. As the executors running at the HPC system would get all needed information in the assigned process records, they become easy to configure and wouldn't need to store credentials locally at the HPC system.

@@

@@

Colonies offer a promising solution for implementing a decentralized and grid-like compute infrastructure, capable of executing Earth Observation (EO) workloads across diverse platforms. By encapsulating computational EO tasks into Colonies processes, interoperability among platforms can be further enhanced. For instance, it becomes possible to create a broker-based architecture that enables OpenEO applications to submit job specifications indirectly to a Colonies server. These specifications are then picked up by executors connecting to different backend providers, e.g. by an executor connected to a Data and Information Access Services (DIAS). Executors can also be leveraged to provision computational resources. For example, automatically provision AWS cloud resources or deploy workflows at HPC systems to handle high-demand EO workloads.

@@

The use of queues also enables a split architecture, where executors can be dynamically managed. In Colonies, it is possible to submit a workflow even if no executor is available to execute the workload. The processes associated with the workflow are simply enqueued until an executor capable of handling the workload is deployed. By designing each executor as a standalone microservice, it becomes easy to scale the system by simply deploying more executors. However, if too few executors are available, a queue of pending processes will begin to build up. This opens up an exciting possibility to delegate colony management responsibility to another software capable of automatically provisioning and deploying executors whenever they are needed, and then also scale down the system to conserve resources and save costs. In line with this, the Cognit project \cite{cognit} aims over the next three years to develop an AI orchestrator capable of automatically managing serverless computing infrastructures.

@@

As a final remark, it is worth highlighting once again the significant advantage of Colonies as a security solution. As a colony is a trusted community of execturs, it can be used to share secrets and credentials among its executor members. By storing credentials, such as one-time passwords or authentication keys in process records, executors can just execute process in run-and-forget style and still be able to access third-party infrastructures without the need to store and manage credentials locally. Additionally, by targeting specific executor types when submitting a function specification, a colony can be further segmented so that only a subset of the executors in a colony can access the credentials, further improving security.

>>>>

Today, utilization of serverless computing is experiencing a significant growth \cite{cognit}, primarily due to its potential to liberate developers from the burden of managing underlying cloud infrastructures. Attempts are currently being made to implement serverless workflow management systems. For example, OpenWolf \cite{openwolf} is a serverless workflow engine designed to utilize the Function-as-a-Service (FaaS) paradigm for composing complex scientific workflows. It is based on OpenFaaS \cite{openfaas}, which allows functions to run on Kubernetes clusters. 
The serverless workflow project \cite{serverlessworkflows} aims at providing a vendor-neutral workflow DSL. Synapse \cite{synapse} is a workflow management system similar to Argo, but that implements the serverless workflow specification. A similar engine is proposed in \cite{scafe}. 
Colonies can also be used to implement serverless workflow management systems. This will be further discussed in Section \ref{faas}. However, in contrast to previous work, Colonies is designed to be platform independent and does therefore not require Kubernetes. By using a zero-trust security protocol, functions can be securely executed by distributed executors deployed anywhere on the Internet. It is important to point out that Colonies does not provide an infrastructure for function execution. Instead, Colonies primary role is to serve as a platform for coordinating function executions which are carried out by distributed executors.

@@

Although microservices have attractive properties and simplify both development and deployment of executors, they also introduce additional complexity that must be managed by the Colonies servers. To implement a workflow management framework supporting distributed microservices, the following challenges must be addressed:

@@

\item \textbf{Service discovery:} The Colonies framework must enable executors to dynamically register and deregister from Colonies servers. It must be possible to deploy executors anywhere on the Internet, even behind firewalls.

@@

The Colonies framework uses a combination of technologies to address the aforementioned challenges. A fundamental design principle is statelessness, which makes Colonies simpler, more reliable and scalable, and easier to implement. To ensure reliability and data consistency, a distributed consensus algorithm is needed to handle high-availability and server crashes. The remainder of this section describes how the Colonies framework is implemented. The next discusses the role of queues to support batch jobs and realtime processing, but also how to implement a loosely coupled system.

@@

Table \ref{proctable} shows an example of a process table. The process table also contains a reference to a function specification depicted in Table \ref{functable}. To assign a process to an executor, the Colonies server searches in the process table to find a process matching a waiting executor. By using the \emph{priority time} column to sort the processes, the process table can function as a queue. This can be done in SQL by utilizing the \emph{order by} clause to sort processes according to their submission time. To make it possible to handle priorities, the submission time is adjusted so that higher priority processes are processed before lower priority processes. When a process is submitted, a \emph{priority time} value is calculated and stored in the process table. Equation \ref{eq:pt} shows how the priority time is calculated for a nanosecond timestamp.

@@

To minimize downtime, Colonies supports high-availability deployments. If one Colonies server crashes, an executor simply need to resend the failed request, which will then be served by another Colonies server replica. However, by introducing multiple Colonies servers, there is again a risk of race conditions when assigning processes to executors. This means that the Colonies must coordinate which replica server should serve incoming assign requests so that precisely one executor is assigned to a particular process.

Raft \cite{raft} is a consensus algorithm specifically designed to manage a replicated log within a distributed system. Raft makes it possible to implement distributed leader elections where a single server takes on the role of leader while the remaining servers act as followers. The leader is responsible for managing the replicated log, processing client requests, and replicating entries to the followers. Followers passively replicate the leader's log and participate in leader elections.

By using the Raft protocol, Colonies can redirect incoming \emph{assign} requests to the leading Colonies server, thereby ensuring that only one Colonies server replica handles such requests. A new leader is automatically elected in the event that a Colonies server replica fails. Figure \ref{fig:ha_deployment} shows an overview of a high-availability Colonies deployment. In addition, the Raft protocol makes it possible to seamlessly update Colonies software by making it possible to upgrade each Colonies server replica individually. Consequently, making Colonies well-suited for Kubernetes deployments.

@@

\subsubsection{Workflows}
A workflow is a series of processes that need to be completed in a specific order. In Colonies, workflows are represented as directed acyclic graphs (DAGs) where nodes represent processes and edges represent dependencies and data flow between processes. Similar to processes, workflows are managed completely stateless. When a workflow is submitted, all processes part of the workflow are submitted and added to the process table similar to how ordinary processes are handled. To control the order processes can be assigned, Colonies sets a flag, \emph{wait for parents} to prevent processes to be assigned to an executor before their dependencies have been satisfied. Note that processes may run in parallel if a sufficient number of executors are available, assuming their parent processes have finished.

@@

When a process is closed, the Colonies server checks its child processes to see if all their parent processes have also completed. If all parents of a child process have finished, the server updates the process table in the database by setting the flag \emph{wait for parents} to false, which allows child processes to be assigned to executors. This operation is also done statelessly when processing the \emph{close} request.

@@

To be able to generate a DAG, an additional database table is required. Table \ref{deptable} shows an example of a dependency table storing relationships between processes. The table is updated when a workflow is submitted to a Colonies server. As mentioned before, when an executor is assigned a process it obtains exclusive rights to that process within the process table. This exclusivity also means that it is possible for an exector to dynamically add new children processes to the workflow DAG on the fly. As a result, it becomes possible to implement MapReduce \cite{mapreduce} data processing patterns similar to Hadoop \cite{hadoop}.

@@

Cron is a time-based job scheduler designed to trigger workflows at predefined intervals. It can for example be used to automate tasks such as routinely fetching data from external systems. To make Colonies robust and easy to implement it is essential that Cron workflows also are managed stateless. This can be achieved by introducing an additional database table to store information about Cron workflow, and let the elected Colonies server leader assume the responsibility of managing Cron workflows through a two-step process.

@@

\subsubsection{Generators}
Another feature of the Colonies framework are so-called generators. Generators are workflow templates that are automatically triggered when specific conditions are met. Generators can be used to accumulate input data into bulks to improve performance or parallelism. For example, an external service downloads satellite images and can use a generator to automatically triggered a workflow when 10 images are collected. This is achieved by sending a special \emph{pack} request containing input data to a Colonies server, targeting a certain generator.

@@

An advantage of using a stateless architecture for workflow management as proposed in the paper is the ability to monitor executor performance to detect any unusual or suspicious behaviors. Because all transactions are recorded in a database, it is possible to track the activity of individual executors and identify any unusual patterns or behaviors. For example, if an executor suddenly starts to consume a lot of resources, it could indicate a performance issue, or it could be a sign of malicious activity, such as a denial-of-service attack. Automatic monitoring can for example be implemented using anomaly detection algorithms and machine learning to grade the performance of each executor. Administrators can then be automatically notified if the grade of an executor exceeds a certain threshold, enabling them to take appropriate actions.

@@

\subsection{Python SDK}
A Colonies application typically consists of a set executors written in different programming languages and deployed across multiple platforms. The executors are small microservices that interact with one another by submitting function specifications or workflows, which are then executed by other executors that are part of the same colony.

The first step in implementing an executor in Python is to create a private key and identity, followed by registering the executor with a colony. Registering an executor can only be carried out by the colony owner with access to the colony private key. Code in Listening \ref{code:regexecutor} illustrates how to register an executor and which functions a specific executor is capable of running. In the example, an executor of type \emph{helloworld\_executor} is registered to run a function named \emph{helloworld}.

@@

\begin{figure}
     \centering
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[scale=0.17]{dashboard1.png}
         \caption{Screenshot of process information.}
     	 \label{fig:dashboard1}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[scale=0.17]{dashboard2.png}
      	 \caption{Screenshot of a workflow.}
     	 \label{fig:dashboard2}
     \end{subfigure}
     \caption{Screenshot of the Colonies Dashboard tool.}
\end{figure}

@@

Figure \ref{fig:dashboard1} shows a screenshot of the Colonies dashboard tool, which enables users to monitor the status of processes and executors. The timeline depicted in the figure allows for real-time monitoring of process execution, similar to tracking a physical package sent through a postal service. As previously discussed, it's also possible to subscribe to a process and receive information about its progress, which can be used as another point of integration. For example, sending a notification on Slack if a process fails.

\subsection{Workflows}
A workflow is just a set of processes with dependencies, which is represented as a DAG internally in Colonies. Workflows can either be submitted using SDKs or the Colonies CLI, which supports submission of JSON files. Listening \ref{code:workflow_spec} shows an example of a workflow consisting of four processes, each requiring different types of executors. The management of workflows is entirely handled by Colonies, and each executor executes only processes that are part of the graph for which it is responsible for. This approach significantly reduces the complexity of designing, implementing and operating large-scale AI applications and data pipelines.

@@

In addition to monitoring processes and executors, the Colonies dashboard can be used to visualize workflows. Figure \ref{fig:dashboard2} shows a DAG created when submitting the workflow code shown in Listing \ref{code:workflow_spec}. This functionality makes it possible to follow workflow execution progress.

@@

\section{Use cases}
Colonies is currently used commerically by a startup company called RockSigma AB\footnote{https://www.rocksigma.com} to implement a compute engine for analysing seismic activities in underground mines. This section will not dig into the details of this specific implementation; instead it will provide an outline of potential future use cases in which Colonies could make a considerable impact.

\subsection{Serverless computing}
\label{faas}
Serverless computing is a cloud-based architecture that allows developers to build and deploy applications without the need to manage the underlying infrastructure. In this model, the cloud service provider dynamically allocates and manages resources such as adding computing power or storage required by an application. Function-as-a-Service (FaaS) is a serverless computing model where developers provide a code snippet that should be executed, and the cloud provider takes care of scaling, configuration and deployment of that code snippet. FaaS is often used in event-driven architectures where functions are triggered by specific events, thus eliminating the need to keep servers continuously running. The Cognit project \cite{cognit} aims to extend the FaaS concept by supporting geographically distributed edge servers in addition to cloud server resources. The project is also going to develop an AI-enabled orchestrator that can decide where to deploy FaaS code.

Colonies offer functionality to call functions in remote executors, a fundamental requirement to implement a FaaS platform. Colonies could also provide service discovery and security isolation to prevent non-authorized function calls. In some cases, it could make sense to manually manage a colony by an administrator, and manually registering and deploying executors. Alternatively, a software solution could be developed to automate the administration and automatically deploy or undeploy executors. Such an approach would be very similar to a Function-as-a-Service (FaaS) platform.

Figure \ref{fig:faas} depicts a Colonies-based FaaS platform. Similar to AWS Lambda or Google Cloud Functions, developers would need to upload code snippets, which would then be packaged and published as Docker containers. To invoke a function, a function specification needs to be submitted to the Colonies server indicating which function to call.

The advantage of Colonies for implementing a FaaS platform is that it introduces an extra level of separation and the possibility to develop a loosely coupled system. For example, if an executor is missing, the process is simply enqueued in the process table. A software agent could then monitor pending processes on the Colonies server, predict the system load, and automatically deploy executors to meet the load, resulting in a highly efficient and scalable system. In the future, the AI orchestrator developed in the Cognit project could also be used together with Colonies to reduce energy consumption or minimize network latency by efficiently utilizing edge servers.

\subsection{HPC integration}
\label{hpc_use_case}
Slurm \cite{slurm} is a workload manager and job scheduler commonly used in HPC environments. In contrast to Colonies, it provides functionality to reserve resources. This is particularly useful when running complex simulations that require a large number of CPUs or a large amount of memory.

@@

@@

Colonies could be used to fill the gap between cloud and HPC and create workflows that can automate processing of data in the cloud while also taking advantage of the vaste computational capacity offered by HPC. For example, a company could develop a workflow that automatically copy training data to a HPC center, train a neural network, move the result back to the cloud, and finally re-deploy a Kubernetes service to use the new model.

Figure \ref{fig:hpc_workflow} shows an example how Colonies could be used in a HPC environment. Several executors run at the HPC system, each with their own responsibilities. The Container Importer Executor is responsible for converting container images to a format that is supported by Slurm, e.g. Singularity containers\footnote{https://docs.sylabs.io}. The Data Importer Executors downloads data to the HPC file system. The Slurm Executor is responsible for submitting and overseeing Slurm jobs, while the Data Exporter executor is responsible for uploading result data to the cloud.

\begin{figure}[h]
	\centering
    \includegraphics[scale=0.43]{hpc_overlay.png}
	\caption{A global grid-like master-control submission system for HPC.}
	\label{fig:hpc_overlay}
\end{figure}

In general, integrating with various authentication services presents a major challenge when developing HPC workflows. For example, the Data Importer may need to access an S3 service to fetch data. The Container Importer must authenticate with a Docker registry to be able to pull a Docker container to the HPC system. In addition, it is not straightforward to connect to the HPC system. In many cases, users have to manually connect using SSH terminals, making it time-consuming and difficult to automate HPC workflows.

With Colonies, users can indirectly control HPC systems by submitting Colonies processes that are subsequently picked up by HPC executors, without requiring direct connection to the HPC system. Essentially, using Colonies as a cloud-based submission system for HPC, connecting different HPC centers together in a grid-like overlay network. Figure \ref{fig:hpc_overlay} depicts such a submission system.

Additionally, the security model provided by Colonies allows different executors to trust one another, thereby enabling the sharing of confidential information, such as S3 secrets. As credentials are stored in the assigned process record, there is no need to store them locally on the HPC system. As a result, HPC executors become more straightforward to configure and deploy, further simplifying creation of HPC workflows.

@@

Colonies could offer a promising solution for implementing a decentralized and grid-like compute infrastructure, capable of executing Earth Observation (EO) workloads across diverse platforms. By encapsulating computational EO tasks into Colonies processes, interoperability among platforms can be further enhanced. For instance, it becomes possible to create a broker-based architecture that enables OpenEO applications to submit job specifications indirectly to a Colonies server. These specifications could then be picked up by executors connected to different backend providers, e.g. a Data and Information Access Services (DIAS) executor. Executors can also be used to leveraged to provision computational resources. For example, automatically provision AWS cloud resources or deploy workflows at HPC systems to handle high-demand EO workloads.

@@

The use of queues also enables a split architecture, where executors can be dynamically managed. In Colonies, it is possible to submit a workflow even if no executor is available to execute the workload. The processes associated with the workflow are simply enqueued until an executor capable of handling the workload is deployed. By designing each executor as a standalone microservice, it becomes easy to scale the system by simply deploying more executors. However, if too few executors are available, a queue of pending processes will begin to build up. This opens up an exciting possibility to delegate colony management responsibility to another software capable of automatically provisioning and deploying executors whenever they are needed, and then also scale down the system to conserve resources and save costs.

@@

As a final remark, it is worth highlighting once again the significant advantage of Colonies as a security solution. As a colony is a trusted community, an executor can share secrets with other colony members. By storing credentials, such as one-time passwords or authentication keys in process records, executors can just execute process in run-and-forget style and be able to access third-party infrastructures without the need to store and manage credentials locally. Additionally, by targeting specific executor types when submitting a function specification, a colony can be further segmented so that only a subset of the executors in a colony can access the credentials, further improving security.
